Estimating duration in the presence of misclassification

### 2024-07-29 biology-data.tex
"data" is plural

The ATACCC and CIS studies have complementary features that make combining information for the studies to estimate the distribution of the duration of detectability an appealing approach.
ATACCC tested individual frequently but had only 57 infections followed for a maximum of 20 days.
The frequent testing makes estimation for these first 20 days reliable, but longer periods are rare and the data does not exist; therefore, the tails of the distribution cannot be estimated.
The CIS had more infrequent testing but a large sample size (thousands of infections) and up to years of follow-up.
Intuitively, information from ATACCC can inform the first two to three weeks of the distribution and information from the CIS can inform the remainder of the distribution.

>>>>

The ATACCC and CIS studies have complementary features that make combining information for the studies to estimate the distribution of the duration of detectability an appealing approach.
ATACCC tested individual frequently but had only 57 infections followed for a maximum of 20 days.
The frequent testing makes estimation for these first 20 days reliable, but longer periods are rare and the data do not exist; therefore, the tails of the distribution cannot be estimated.
The CIS had more infrequent testing but a large sample size (thousands of infections) and up to years of follow-up.
Intuitively, information from ATACCC can inform the first two to three weeks of the distribution and information from the CIS can inform the remainder of the distribution.

### 2024-07-29 cis-perfect-testing.tex
Christl's first comment on simulations
"data" is plural

In \cref{perf-test:sec:problem}, I describe the challenges analysing the data presents.
This motivates using a survival analysis framework which I introduce in \cref{perf-test:sec:surv-analysis}, before deriving the statistical model for the CIS setting in \cref{perf-test:sec:model}.
In \cref{perf-test:sec:parameters-priors}, I consider how to parameterize the survival function and choose its priors, including vague and smoothing priors as well as a prior constructed to include the information from \cref{E-ATACCC}.
I evaluate the model and priors using a simulation study (\cref{perf-test:sec:simulation-study}), showing that they successfully recover $S(t)$.
Finally, I discuss the results in \cref{perf-test:sec:discussion}.

@@

To form the duration distribution used in the simulation, we combine the two estimates.
The pdf over the first 30 days is proportional to the ATACCC estimate, with the rest proportional to this CIS-based estimate.
Denote by $f_A(t)$ the distribution function estimated in \cref{E-ATACCC} and $f_C(t)$ that from the CIS-based estimates just derived.
Then define:
$$
f_S'(t) = \begin{cases}
	f_A(t) &t \leq 30 \\
	f_C(t) &t > 30
\end{cases}
$$
Then the distribution used in the simulation is the normalised version of this: $f_S(t) = f'_S(t)/\sum_i f_S'(i)$.
These curves and the combined curve are compared in \cref{perf-test:fig:duration-dist}.
Individual $i$'s duration of detectability, $D_i$, is then an independent draw from this distribution.
\begin{figure}
  \centering \includegraphics{cis-perfect-testing/input-duration-dists}
  \caption[Comparison of duration distributions]{Comparison of the duration distribution used for the simulation. ATACCC is the posterior mean from \cref{E-ATACCC}'s analysis, CIS is Sarah Walker's analysis, and Combined is the combination of these used for the simulation. \label{perf-test:fig:duration-dist}}
\end{figure}

>>>>

In \cref{perf-test:sec:problem}, I describe the challenges presented by analysing the data.
This motivates using a survival analysis framework which I introduce in \cref{perf-test:sec:surv-analysis}, before deriving the statistical model for the CIS setting in \cref{perf-test:sec:model}.
In \cref{perf-test:sec:parameters-priors}, I consider how to parameterize the survival function and choose its priors, including vague and smoothing priors as well as a prior constructed to include the information from \cref{E-ATACCC}.
I evaluate the model and priors using a simulation study (\cref{perf-test:sec:simulation-study}), showing that they successfully recover $S(t)$.
Finally, I discuss the results in \cref{perf-test:sec:discussion}.

@@

To form the duration distribution used in the simulation, we combine the two estimates.
The pdf over the first 30 days is proportional to the ATACCC estimate, with the rest proportional to this CIS-based estimate.
Denote by $f_A(t)$ the distribution function estimated in \cref{E-ATACCC} and $f_C(t)$ that from the CIS-based estimates just derived.
Then define:
$$
f_S'(t) = \begin{cases}
	f_A(t) &t \leq 30 \\
	f_C(t) &t > 30
\end{cases}
$$
Then the distribution used in the simulation is the normalised version of this: $f_S(t) = f'_S(t)/\sum_i f_S'(i)$.
These curves and the combined curve are compared in \cref{perf-test:fig:duration-dist}.
Compared to the ATACCC estimate alone, incorporating the CIS portion shifts some probability mass into the tail of the distribution.
Therefore, the survival probability is initially higher but later lower than ATACCC alone.
Individual $i$'s duration of detectability, $D_i$, is then an independent draw from this distribution.
\begin{figure}
  \centering \includegraphics{cis-perfect-testing/input-duration-dists}
  \caption[Comparison of duration distributions]{Comparison of the duration distribution used for the simulation. ATACCC is the posterior mean from \cref{E-ATACCC}'s analysis, CIS is Sarah Walker's analysis, and Combined is the combination of these used for the simulation. \label{perf-test:fig:duration-dist}}
\end{figure}

### 2024-07-29 introduction.tex
"data" is plural

An example of incidence-type data are the number of cases identified through community testing, which relies on individuals presenting for a test.
Individuals' ability and willingness to be tested varies over time, which needs to be accounted for in order to perform unbiased inference.

@@

However, severe consequences can be rare in some population subgroups (\eg young, otherwise healthy individuals infected with SARS-CoV-2~\autocite{wardSero,bhopalChildren}), meaning they rarely appear in this data.
Therefore, the data does not provide much information on incidence in those subgroups.
Furthermore, the delay between infection and the severe event occurring means that recent infections cannot be estimated with much precision~\autocite{swallow2022challenges}.

@@

Where possible, aggregated, non-disclosive versions of the datasets are available alongside the code.
The original ATACCC data used is available alongside the code.
The original CIS data can only be accessed by accredited researchers within the ONS's SRS (Secure Research Service); see \url{https://www.ons.gov.uk/aboutus/whatwedo/statistics/requestingstatistics/secureresearchservice} for details.

>>>>

An example of incidence-type data is the number of cases identified through community testing, which relies on individuals presenting for a test.
Individuals' ability and willingness to be tested varies over time, which needs to be accounted for in order to perform unbiased inference.

@@

However, severe consequences can be rare in some population subgroups (\eg young, otherwise healthy individuals infected with SARS-CoV-2~\autocite{wardSero,bhopalChildren}), meaning they rarely appear in these data.
Therefore, the data do not provide much information on incidence in those subgroups.
Furthermore, the delay between infection and the severe event occurring means that recent infections cannot be estimated with much precision~\autocite{swallow2022challenges}.

@@

Where possible, aggregated, non-disclosive versions of the datasets are available alongside the code.
The original ATACCC data used are available alongside the code.
The original CIS data can only be accessed by accredited researchers within the ONS's SRS (Secure Research Service); see \url{https://www.ons.gov.uk/aboutus/whatwedo/statistics/requestingstatistics/secureresearchservice} for details.

### 2024-07-29 transmission.tex
"data" is plural

In this set-up, the parameters $i^+$, $q_c$, and $\psi$ are fully identified; $\pi_0$ and $\rho$ are not identified at all; the other parameters are partially identified, their posterior has moved away from the prior but there is insufficient information in the data to fully overcome the prior (see \cref{SEIR:fig:true-vs-posterior}).
The ability to identify $q_c$ is an advantage of using data which includes measurements on children, unlike data on hospitalisations or deaths.

@@

The increases in $\beta_w$ in December coincide with the rise of the Alpha variant~\autocite{lythgoeLineage}, known to be more transmissible~\autocite[e.g.][]{daviesEstimated}.
A multiple variant model might be able to capture this rise more accurately.
Such a model would stratify positive tests in the CIS by pre-Alpha or Alpha variant.
Then, each variant would be explicitly included in the model to fit to this data.

>>>>

In this set-up, the parameters $i^+$, $q_c$, and $\psi$ are fully identified; $\pi_0$ and $\rho$ are not identified at all; the other parameters are partially identified, their posterior has moved away from the prior but there is insufficient information in the data to fully overcome the prior (see \cref{SEIR:fig:true-vs-posterior}).
The ability to identify $q_c$ is an advantage of using data which include measurements on children, unlike data on hospitalisations or deaths.

@@

The increases in $\beta_w$ in December coincide with the rise of the Alpha variant~\autocite{lythgoeLineage}, known to be more transmissible~\autocite[e.g.][]{daviesEstimated}.
A multiple variant model might be able to capture this rise more accurately.
Such a model would stratify positive tests in the CIS by pre-Alpha or Alpha variant.
Then, each variant would be explicitly included in the model to fit to these data.
